Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-27 article.tex
fix: network density
typo

We report here experiments on the \textit{normal-1,2,3}, and \textit{4}
datasets provided by the organizers of the Connectomics
challenge. Each of these datasets is obtained from the simulation
\citep{stetter2012model} of different neural networks of 1000 neurons
and about 15000 edges each (i.e., a network density of about
1.5\%). Each neuron is described by a calcium fluorescence time series
of length $T=179500$. All inference methods compared here provide a
ranking of all pairs of neurons according to some association
score. To assess the quality of this ranking, we compute both ROC and
precision-recall curves against the ground truth network, which are
summarized using the area under the curve and respectively  denoted AUROC
and AUPRC. Only the AUROC score was used to rank the challenge
participants but the precision-recall curve has been shown to be a
more sensible metric for network inference, especially when network
density is small (see e.g., \cite{schrynemackers2013protocols}).
Since neurons are not self-connected in the ground truth networks
(i.e., $(i, i) \not \in E, \forall i \in V$), we have manually set the
score of such edges to the minimal possible association score before
computing ROC and PR curves.

@@

In this paper, we outlined a simple but efficient methodology for the problem
of connectome inference from calcium imaging data. Our approach is made of two
steps: (i) processing fluorescence data to detect neural peak activities and
(ii) inferring the degree of association between neurons from partial
correlation statistics. Its simplified variant shows to outperform other
network inference methods while its optimized version proved to be best method
on the Connectomics challenge. Given its simplicity and good performance, we
therefore believe that the (simplified) methodology presented in this work
would constitute a solid and easily reproducible baseline for further work in
the field of connectome inference.

>>>>

We report here experiments on the \textit{normal-1,2,3}, and \textit{4}
datasets provided by the organizers of the Connectomics
challenge. Each of these datasets is obtained from the simulation
\citep{stetter2012model} of different neural networks of 1000 neurons
and about 12000 edges each (i.e., a network density of about
1.2\%). Each neuron is described by a calcium fluorescence time series
of length $T=179500$. All inference methods compared here provide a
ranking of all pairs of neurons according to some association
score. To assess the quality of this ranking, we compute both ROC and
precision-recall curves against the ground truth network, which are
summarized using the area under the curve and respectively  denoted AUROC
and AUPRC. Only the AUROC score was used to rank the challenge
participants but the precision-recall curve has been shown to be a
more sensible metric for network inference, especially when network
density is small (see e.g., \cite{schrynemackers2013protocols}).
Since neurons are not self-connected in the ground truth networks
(i.e., $(i, i) \not \in E, \forall i \in V$), we have manually set the
score of such edges to the minimal possible association score before
computing ROC and PR curves.

@@

In this paper, we outlined a simple but efficient methodology for the problem
of connectome inference from calcium imaging data. Our approach is made of two
steps: (i) processing fluorescence data to detect neural peak activities and
(ii) inferring the degree of association between neurons from partial
correlation statistics. Its simplified variant shows to outperform other
network inference methods while its optimized version proved to be the best method
on the Connectomics challenge. Given its simplicity and good performance, we
therefore believe that the (simplified) methodology presented in this work
would constitute a solid and easily reproducible baseline for further work in
the field of connectome inference.
